---
title: "Data Visualization with Plotly"
author: "Hesa"
date: "March 18, 2025"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(dplyr)
```

## Introduction

- This presentation demonstrates the use of Plotly for creating interactive data visualizations in R
- Plotly allows for creation of publication-quality charts with interactive features
- Features include tooltips, zooming, and panning
- Created on March 18, 2025

## Interactive Data Visualization {.flexbox .vcenter}

```{r plotly-visualization, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
# Sample temperature anomaly data
years <- 1880:2024
set.seed(123) # For reproducibility
temp_anomalies <- cumsum(rnorm(length(years), mean=0.01, sd=0.1))

# Create data frame
climate_data <- data.frame(
  Year = years,
  TempAnomaly = temp_anomalies
)

# Create color gradient based on temperature values
colors <- colorRampPalette(c("blue", "lightblue", "yellow", "orange", "red"))(100)
color_indices <- cut(climate_data$TempAnomaly, breaks=100, labels=FALSE)

# Create the Plotly figure
p <- plot_ly(climate_data, x = ~Year, y = ~TempAnomaly, 
             type = 'scatter', mode = 'lines+markers',
             marker = list(size = 5, color = colors[color_indices]),
             line = list(color = 'rgba(100, 100, 100, 0.5)'),
             hoverinfo = 'text',
             text = ~paste("Year:", Year, "<br>Temperature Anomaly:", round(TempAnomaly, 2), "째C")) %>%
  layout(title = "Global Temperature Anomalies (1880-2024)",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Temperature Anomaly (째C)"),
         hovermode = "closest")

p
```

## Code Behind the Visualization

```{r code-example, eval=FALSE}
# Create the Plotly figure
p <- plot_ly(climate_data, x = ~Year, y = ~TempAnomaly, 
             type = 'scatter', mode = 'lines+markers',
             marker = list(size = 5, color = colors[color_indices]),
             line = list(color = 'rgba(100, 100, 100, 0.5)'),
             hoverinfo = 'text',
             text = ~paste("Year:", Year, 
                          "<br>Temperature Anomaly:", 
                          round(TempAnomaly, 2), "째C")) %>%
  layout(title = "Global Temperature Anomalies (1880-2024)",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Temperature Anomaly (째C)"),
         hovermode = "closest")
```

## Analysis

Key observations:

- Clear upward trend in temperature anomalies
- Rate of increase appears to accelerate in recent decades
- Year-to-year variability is significant
- Long-term warming trend is unmistakable

## Interactive Features

This Plotly visualization includes:

- **Hover tooltips**: Move your cursor over data points
- **Zooming**: Click and drag to zoom
- **Panning**: After zooming, click and drag to move
- **Download**: Use camera icon to download the plot
- **Reset**: Double-click to reset the view

## Alternative Visualization: 3D Surface Plot

```{r surface-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
# Create sample 3D data
x_seq <- seq(-5, 5, length.out = 50)
y_seq <- seq(-5, 5, length.out = 50)
z_matrix <- outer(x_seq, y_seq, function(x, y) sin(sqrt(x^2 + y^2)))

# Create surface plot
surface_plot <- plot_ly(x = x_seq, y = y_seq, z = z_matrix, type = "surface",
                     colorscale = "Viridis") %>%
  layout(title = "3D Surface Plot Example",
         scene = list(
           xaxis = list(title = "X"),
           yaxis = list(title = "Y"),
           zaxis = list(title = "Z")
         ))

surface_plot
```

## Conclusion

- Plotly provides powerful tools for interactive data visualization
- Enhances data exploration and communication
- Interactive elements improve audience engagement
- R + Plotly + R Markdown makes presentation creation smooth

## Thank You!

Created on: March 18, 2025
